---
title: "hw1_p8105"
author: "Brandon Rojas"
date: "9/29/2021"
output: github_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
```

## Problem 1
* Create a data frame comprised of:
    + a random sample of size 10 from a standard Normal distribution 
    + a logical vector indicating whether elements of the sample are greater than 0
    + a character vector of length 10 
    + a factor vector of length 10, with 3 different factor “levels”

When taking the mean of each column in the dataframe, we notice that norm_samp, and "norm_samp_pos" return numeric values, as these logical values are treated as 1/0, and norm_samp is a series of numeric values.

factor_vector and char_vec are not numeric or logical, and are factors and character vectors. Characters cannot be added or divided, thus it returns "NA", the same goes for factors.

We can coerce these into being numeric as follows:

as.numeric(r_samp_tbl[["norm_samp_pos"]])

as.numeric(r_samp_tbl[["char_vec"]])

as.numeric(r_samp_tbl[["factor_vector"]])

This will convert the logical into true numerical values 0 & 1, the characters become NAs as they have no valid conversion, and the factor vectors are receive values 1 to 3 because 3 levels of factor, "high, medium and low" exist. Noted that coercing the factors into numeric would allow you to take the mean.

```{r problem_1}
set.seed(1337)

r_samp_tbl = tibble(
  norm_samp = rnorm(10),                       # random sample size 10, std normal~
  factor_vector = (factor(c("high", "medium", "low", "high", "medium", "low", "high", "medium", "low", "high"))),
  norm_samp_pos = norm_samp > 0,                                 # elements greater than zero
  char_vec = character(10),                                      #char vec of length 10
)

r_samp_tbl

```

## Taking Means
```{r problem_1_means}
##Means
mean(r_samp_tbl[["norm_samp"]])
mean(r_samp_tbl[["factor_vector"]])
mean(r_samp_tbl[["norm_samp_pos"]])
mean(r_samp_tbl[["char_vec"]])
```


## Problem 2

```{r penguin_load}
data("penguins", package = "palmerpenguins")
```

The Palmer Penguin data, including mean flipper length, and the observation types with the number of recorded values can be summarized as follows:

```{r penguins}
summary(penguins)

print(paste("Number of Columns:", ncol(penguins)))
print(paste("Number of Rows:", nrow(penguins)))
      
```

Visualized:
```{r penguin_chart}
# from http://www.cookbook-r.com/Graphs/Colors_(ggplot2)/#a-colorblind-friendly-palette
color_blind_palette <- c("#999999", "#E69F00", "#56B4E9", "#009E73", "#F0E442", "#0072B2", "#D55E00", "#CC79A7") 


ggplot(penguins, 
       aes(x = bill_length_mm, y = flipper_length_mm, color = species)) + 
       scale_color_manual(values= color_blind_palette) + 
       geom_point(size = 1.5) + 
       ggtitle("Anatomical variance of Palmer Archipelago Penguins") + 
       xlab("bill length (mm)") + 
       ylab("flipper length (mm)")

ggsave("br2598_hw1_p8105_scatterplot.png")
```



